---
title: "lesson5_problem_set"
output: html_document
---

```{r diamonds data set}
library(ggplot2)
data("diamonds")
?diamonds
```

## Price Histogram with Facet  
# Create a histogram of diamond prices.
# Facet the histogram by diamond color
# and use cut to color the histogram bars.

# The plot should look something like this.
# http://i.imgur.com/b5xyrOu.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the histogram using
# scale_fill_brewer(type = 'qual')

```{r price histogram with facet}
ggplot(aes(x=price, fill=cut), data=diamonds) +
  geom_histogram() +
  scale_x_log10() +
  facet_wrap(~ color)
```

## Price vs. table colored by cut

# Create a scatterplot of diamond price vs.
# table and color the points by the cut of
# the diamond.

# The plot should look something like this.
# http://i.imgur.com/rQF9jQr.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the scatterplot using
# scale_color_brewer(type = 'qual')

```{r price vs. table colored by cut}
ggplot(aes(x=table, y=price), data=diamonds) +
  scale_x_continuous(limits = c(50, 80), breaks = seq(50, 80, 2)) +
  geom_point(aes(color = cut)) +
  scale_color_brewer(type = 'qua')

```

## Typical Table Value  
# What is the typical table range for the majority of diamonds of ideal cut?
# 53 to 57
# What is the typical table range for the majority of diamonds of premium cut?
# 57 to 62

## Price vs. Volume and Diamond  

# Create a scatterplot of diamond price vs.
# volume (x * y * z) and color the points by
# the clarity of diamonds. Use scale on the y-axis
# to take the log10 of price. You should also
# omit the top 1% of diamond volumes from the plot.

# Note: Volume is a very rough approximation of
# a diamond's actual volume.

# The plot should look something like this.
# http://i.imgur.com/excUpea.jpg

# Note: In the link, a color palette of type
# 'div' was used to color the scatterplot using
# scale_color_brewer(type = 'div')

```{r Price vs. Volume and Diamond}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(subset(diamonds, volume <= quantile(volume, 0.99) & volume > 0), aes(x=volume, y=price, color= clarity)) +
  geom_point() +
  scale_y_log10() +
  scale_color_brewer(type = 'div') +
  theme_minimal()


```

## Proportion of Friendship Initiated  

# Many interesting variables are derived from two or more others.
# For example, we might wonder how much of a person's network on
# a service like Facebook the user actively initiated. Two users
# with the same degree (or number of friends) might be very
# different if one initiated most of those connections on the
# service, while the other initiated very few. So it could be
# useful to consider this proportion of existing friendships that
# the user initiated. This might be a good predictor of how active
# a user is compared with their peers, or other traits, such as
# personality (i.e., is this person an extrovert?).

# Your task is to create a new variable called 'prop_initiated'
# in the Pseudo-Facebook data set. The variable should contain
# the proportion of friendships that the user initiated.

```{r proportion of Friendship Initiated}
pf <- read.csv('../pseudo_facebook.tsv', sep='\t')
library('dplyr')
```

```{r}
pf$prop_initiated <- pf$friendships_initiated / pf$friend_count
```


## Proportion initiated vs. tenure 

# Create a line graph of the median proportion of
# friendships initiated ('prop_initiated') vs.
# tenure and color the line segment by
# year_joined.bucket.

# Recall, we created year_joined.bucket in Lesson 5
# by first creating year_joined from the variable tenure.
# Then, we used the cut function on year_joined to create
# four bins or cohorts of users.

# (2004, 2009]
# (2009, 2011]
# (2011, 2012]
# (2012, 2014]

# The plot should look something like this.
# http://i.imgur.com/vNjPtDh.jpg
# OR this
# http://i.imgur.com/IBN1ufQ.jpg


```{r proportion initiated vs. tenure}
pf$year_joined <- floor(2014 - pf$tenure / 365)
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))
table(pf$year_joined.bucket)
```

```{r}
pf <- pf %>% mutate(prop_initiated = ifelse(friend_count >0, friendships_initiated/friend_count, 0))

```

```{r}
ggplot(aes(x=tenure, y=prop_initiated, colour=year_joined.bucket), data=subset(pf, tenure>0 & !is.na(year_joined.bucket))) +
  geom_line(stat='summary', fun.y=median) +
  theme_minimal()

```

## Smoothing Prop_initiated  

# Smooth the last plot you created of
# of prop_initiated vs tenure colored by
# year_joined.bucket. You can bin together ranges
# of tenure or add a smoother to the plot.

# There won't be a solution image for this exercise.
# You will answer some questions about your plot in
# the next two exercises.

```{r smoothing prop_initiated}
ggplot(aes(x=tenure, y=prop_initiated, colour=year_joined.bucket), data=subset(pf, tenure>0 & !is.na(year_joined.bucket))) +
  geom_smooth(stat='summary', fun.y=median) +
  theme_minimal()
```

```{r smoothing prop_initiated}
ggplot(aes(x=tenure, y=prop_initiated, colour=year_joined.bucket), data=subset(pf, tenure>0 & !is.na(year_joined.bucket))) +
  geom_smooth() +
  theme_minimal()
```

## Greatest prop_initiated
## On average, which group initiated the greatest
## poportion of its Facebook friendships? The plot with
## the smoother that you created in the last exercise can
## help you answer this question.

## People who joined after 2012.

```{r Greatest prop_initiated}

```
## Largest group mean prop_initiated

## For the group with the largest proportion of 
## friendships initated, what is the group's average 
## (mean) proportion on friendships initiated?

```{r prop_initiated vs. tenure}
with(subset(pf, year_joined.bucket == '(2012,2014]'), mean(prop_initiated))
```

## Price / carat binned,  faceted

# Create a scatter plot of the price/carat ratio
# of diamonds. The variable x should be
# assigned to cut. The points should be colored
# by diamond color, and the plot should be
# faceted by clarity.

# The plot should look something like this.
# http://i.imgur.com/YzbWkHT.jpg.

# Note: In the link, a color palette of type
# 'div' was used to color the histogram using
# scale_color_brewer(type = 'div')

```{r price / carat binned, faceted}
ggplot(aes(x=cut, y=price/carat, color=color), data=diamonds) +
  geom_jitter() +
  facet_wrap(~ clarity) +
  scale_color_brewer(type = 'div') 

```

## Gapminder Multivariable  


```{r Gapminder Multivariable}

```
