---
title: "lesson5_problem_set"
output: html_document
---

```{r diamonds data set}
library(ggplot2)
data("diamonds")
?diamonds
```


## Price Histogram with Facet  
# Create a histogram of diamond prices.
# Facet the histogram by diamond color
# and use cut to color the histogram bars.

# The plot should look something like this.
# http://i.imgur.com/b5xyrOu.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the histogram using
# scale_fill_brewer(type = 'qual')

```{r price histogram with facet}
ggplot(aes(x=price, fill=cut), data=diamonds) +
  geom_histogram() +
  scale_x_log10() +
  facet_wrap(~ color)
```

## Price vs. table colored by cut

# Create a scatterplot of diamond price vs.
# table and color the points by the cut of
# the diamond.

# The plot should look something like this.
# http://i.imgur.com/rQF9jQr.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the scatterplot using
# scale_color_brewer(type = 'qual')

```{r price vs. table colored by cut}
ggplot(aes(x=table, y=price), data=diamonds) +
  scale_x_continuous(limits = c(50, 80), breaks = seq(50, 80, 2)) +
  geom_point(aes(color = cut)) +
  scale_color_brewer(type = 'qua')

```

## Typical Table Value  
# What is the typical table range for the majority of diamonds of ideal cut?
# 53 to 57
# What is the typical table range for the majority of diamonds of premium cut?
# 57 to 62

## Price vs. Volume and Diamond  

# Create a scatterplot of diamond price vs.
# volume (x * y * z) and color the points by
# the clarity of diamonds. Use scale on the y-axis
# to take the log10 of price. You should also
# omit the top 1% of diamond volumes from the plot.

# Note: Volume is a very rough approximation of
# a diamond's actual volume.

# The plot should look something like this.
# http://i.imgur.com/excUpea.jpg

# Note: In the link, a color palette of type
# 'div' was used to color the scatterplot using
# scale_color_brewer(type = 'div')

```{r Price vs. Volume and Diamond}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(subset(diamonds, volume <= quantile(volume, 0.99) & volume > 0), aes(x=volume, y=price, color= clarity)) +
  geom_point() +
  scale_y_log10() +
  scale_color_brewer(type = 'div') +
  theme_minimal()


```

## Proportion of Friendship Initiated  

# Many interesting variables are derived from two or more others.
# For example, we might wonder how much of a person's network on
# a service like Facebook the user actively initiated. Two users
# with the same degree (or number of friends) might be very
# different if one initiated most of those connections on the
# service, while the other initiated very few. So it could be
# useful to consider this proportion of existing friendships that
# the user initiated. This might be a good predictor of how active
# a user is compared with their peers, or other traits, such as
# personality (i.e., is this person an extrovert?).

# Your task is to create a new variable called 'prop_initiated'
# in the Pseudo-Facebook data set. The variable should contain
# the proportion of friendships that the user initiated.

```{r proportion of Friendship Initiated}
pf <- read.csv('../pseudo_facebook.tsv', sep='\t')
library('dplyr')
```

```{r}
pf$prop_initiated <- pf$friendships_initiated / pf$friend_count
```


## Proportion initiated vs. tenure 

# Create a line graph of the median proportion of
# friendships initiated ('prop_initiated') vs.
# tenure and color the line segment by
# year_joined.bucket.

# Recall, we created year_joined.bucket in Lesson 5
# by first creating year_joined from the variable tenure.
# Then, we used the cut function on year_joined to create
# four bins or cohorts of users.

# (2004, 2009]
# (2009, 2011]
# (2011, 2012]
# (2012, 2014]

# The plot should look something like this.
# http://i.imgur.com/vNjPtDh.jpg
# OR this
# http://i.imgur.com/IBN1ufQ.jpg


```{r proportion initiated vs. tenure}
pf$year_joined <- floor(2014 - pf$tenure / 365)
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))
table(pf$year_joined.bucket)
```

```{r}
pf <- pf %>% mutate(prop_initiated = ifelse(friend_count >0, friendships_initiated/friend_count, 0))

```

```{r}
ggplot(aes(x=tenure, y=prop_initiated, colour=year_joined.bucket), data=subset(pf, tenure>0 & !is.na(year_joined.bucket))) +
  geom_line(stat='summary', fun.y=median) +
  theme_minimal()

```

## Smoothing Prop_initiated  

# Smooth the last plot you created of
# of prop_initiated vs tenure colored by
# year_joined.bucket. You can bin together ranges
# of tenure or add a smoother to the plot.

# There won't be a solution image for this exercise.
# You will answer some questions about your plot in
# the next two exercises.

```{r smoothing prop_initiated}
ggplot(aes(x=tenure, y=prop_initiated, colour=year_joined.bucket), data=subset(pf, tenure>0 & !is.na(year_joined.bucket))) +
  geom_smooth(stat='summary', fun.y=median) +
  theme_minimal()
```

```{r smoothing prop_initiated}
ggplot(aes(x=tenure, y=prop_initiated, colour=year_joined.bucket), data=subset(pf, tenure>0 & !is.na(year_joined.bucket))) +
  geom_smooth() +
  theme_minimal()
```

## Greatest prop_initiated
## On average, which group initiated the greatest
## poportion of its Facebook friendships? The plot with
## the smoother that you created in the last exercise can
## help you answer this question.

## People who joined after 2012.

```{r Greatest prop_initiated}

```
## Largest group mean prop_initiated

## For the group with the largest proportion of 
## friendships initated, what is the group's average 
## (mean) proportion on friendships initiated?

```{r prop_initiated vs. tenure}
with(subset(pf, year_joined.bucket == '(2012,2014]'), mean(prop_initiated))
```

## Price / carat binned,  faceted

# Create a scatter plot of the price/carat ratio
# of diamonds. The variable x should be
# assigned to cut. The points should be colored
# by diamond color, and the plot should be
# faceted by clarity.

# The plot should look something like this.
# http://i.imgur.com/YzbWkHT.jpg.

# Note: In the link, a color palette of type
# 'div' was used to color the histogram using
# scale_color_brewer(type = 'div')

```{r price / carat binned, faceted}
ggplot(aes(x=cut, y=price/carat, color=color), data=diamonds) +
  geom_jitter() +
  facet_wrap(~ clarity) +
  scale_color_brewer(type = 'div') 

```

## Gapminder Multivariable  

```{r}
install.packages('dplyr')
library(dplyr)
install.packages('tidyr')
library(tidyr)
install.packages("gridExtra")
library("gridExtra")
```

# I chose to investigate foreign direct investment inflow (% of GDP),  outflow (% of GDP), and investment (% of GDP) from gapminder.  I did a comparison study between US and China.

```{r Gapminder Multivariable}
gdp_per_capital_growth <- read.csv('./Gapminder_project/GDPpercapitaconstant2000US/data.csv', sep = ',', head = TRUE)
gdp_per_capital <- read.csv('./Gapminder_project/indicatorwdigdp_percapita_growth/data.csv', sep = ',', head = TRUE)
gdp_per_employee <- read.csv('./Gapminder_project/indicator_gdp_per_employee/data.csv', sep = ',', head = TRUE)
gdp_per_working_hour<- read.csv('./Gapminder_project/indicator_gdp_per_working_hour/data.csv', sep = ',', head = TRUE)

agriculture<- read.csv('./Gapminder_project/agriculture_p_of_GDP/data.csv', sep = ',', head = TRUE)
industry<- read.csv('./Gapminder_project/industry_p_of_GDP/data.csv', sep = ',', head = TRUE)

foreign_inv_inflow<- read.csv('./Gapminder_project/foreign_investment_inflow/data.csv', sep = ',', head = TRUE)
foreign_inv_outflow<- read.csv('./Gapminder_project/foreign_investment_outflow/data.csv', sep = ',', head = TRUE)
investment<- read.csv('./Gapminder_project/investment/data.csv', sep = ',', head = TRUE)

export<- read.csv('./Gapminder_project/exports_p_of_GDP/data.csv', sep = ',', head = TRUE)

```

```{r}
colnames(gdp_per_capital_growth)[1] <- 'country'
colnames(gdp_per_capital)[1] <- 'country'
colnames(gdp_per_employee)[1] <- 'country'
colnames(gdp_per_working_hour)[1] <- 'country'

per_capital_growth <- gdp_per_capital_growth %>%
  gather('year', 'per_capital_growth', 2:53)
per_capital_growth$year <- gsub('X', '', per_capital_growth$year)

per_capital <- gdp_per_capital %>%
  gather('year', 'per_capital', 2:53)
per_capital$year <- gsub('X', '', per_capital$year)

per_employee <- gdp_per_employee %>%
  gather('year', 'per_employee', 2:29)
per_employee$year <- gsub('X', '', per_employee$year)

per_working_hour <- gdp_per_working_hour %>%
  gather('year', 'per_working_hour', 2:29)
per_working_hour$year <- gsub('X', '', per_working_hour$year)

```

```{r us GDP per factors (capital, capital growth, employee, working hour)}
us_gdp_factors <- left_join(subset(per_capital, country=='United States'), subset(per_capital_growth, country=='United States'), by = 'year')
us_gdp_factors <- left_join(us_gdp_factors, subset(per_employee, country=='United States'))
us_gdp_factors <- left_join(us_gdp_factors, subset(per_working_hour, country=='United States'))
```

```{r}
cor.test(us_gdp_factors$per_capital, us_gdp_factors$per_capital_growth)
cor.test(us_gdp_factors$per_capital, us_gdp_factors$per_employee)
cor.test(us_gdp_factors$per_capital, us_gdp_factors$per_working_hour)
cor.test(us_gdp_factors$per_capital_growth, us_gdp_factors$per_employee)
cor.test(us_gdp_factors$per_capital_growth, us_gdp_factors$per_working_hour)

```
# -0.2505442, 0.118675, 0.09153776, 0.9957948, 0.9860283, 
```{r}
summary(us_gdp_factors$per_capital)
us_gdp_factors$per_capital.bucket <- cut(us_gdp_factors$per_capital, breaks = c(-5, 0, 2, 7))
```

```{r}
p1 <- ggplot(aes(x=year, y=per_capital_growth), data=us_gdp_factors) + 
  geom_point(aes(color=per_capital.bucket)) 
  
p2 <- ggplot(aes(x=year, y=per_working_hour), data=us_gdp_factors) + 
  geom_point(aes(color=per_capital.bucket)) 

p3 <- ggplot(aes(x=year, y=per_employee), data=us_gdp_factors) + 
  geom_point(aes(color=per_capital.bucket)) 

grid.arrange(p1, p2, p3)

ggsave('us_GDP_per_factors.png')

```
# For US, GDP per capital growth but not GDP per capital is highly coorelated to GDP per employee and GDP per working hour.  That means GDP per capital growth manifast through employees work and working hour.  
# GDP per employee is highly coorelated to GDP per working hour.
# If GDP per capital is negative, GDP per capital growth would stay flat or decrease.

```{r - investment, foreign investment inflow, foreign investment outflow}
colnames(foreign_inv_inflow)[1] <- 'country'
colnames(foreign_inv_outflow)[1] <- 'country'
colnames(investment)[1] <- 'country'

foreign_inv_outflow<- foreign_inv_outflow %>%
  gather('year', 'foreign_outflow', 2:53)
foreign_inv_outflow$year <- gsub('X', '', foreign_inv_outflow$year)

foreign_inv_inflow<- foreign_inv_inflow %>%
  gather('year', 'foreign_inflow', 2:43)
foreign_inv_inflow$year <- gsub('X', '', foreign_inv_inflow$year)

investment<- investment %>%
  gather('year', 'investment', 2:53)
investment$year <- gsub('X', '', investment$year)

us_investment_foreign <- left_join(subset(foreign_inv_inflow, country=='United States'), subset(foreign_inv_outflow, country=='United States'), by = 'year')
us_investment_foreign <- left_join(us_investment_foreign, subset(investment, country=='United States'), by = 'year')

china_investment_foreign <- left_join(subset(foreign_inv_inflow, country=='China'), subset(foreign_inv_outflow, country=='China'), by = 'year')
china_investment_foreign <- left_join(china_investment_foreign, subset(investment, country=='China'), by = 'year')

investment_and_foreign_IO <- bind_rows(us_investment_foreign, china_investment_foreign)

investment_and_foreign_IO[c("foreign_outflow", "foreign_inflow")][is.na(investment_and_foreign_IO[c("foreign_outflow", "foreign_inflow")])] <- 0

investment_and_foreign_IO <- subset(investment_and_foreign_IO, select = -c(country.x, country.y))

investment_and_foreign_IO$net_in <- as.numeric(us_investment_foreign$foreign__inflow) - as.numeric(us_investment_foreign$foreign_outflow)

investment_and_foreign_IO <- investment_and_foreign_IO %>%
  mutate(net_in = as.numeric(foreign_inflow) - as.numeric(foreign_outflow))
  
```

```{r}
ggplot(aes(x=year, y=investment, fill=country), data=investment_and_foreign_IO) +
  geom_bar(stat='identity')
ggsave('us_and_china_investment_comparision.png')
```
# China has higher level of investment than US, double US amount in general.

```{r - US and China net inflow comparison plot}
inflow <- ggplot(aes(x = year, y=foreign_inflow, fill=country), data=investment_and_foreign_IO) + 
  geom_histogram(stat = 'identity') +
  geom_point(aes(y=investment,color=country))

outflow <- ggplot(aes(x = year, y=foreign_outflow, fill=country), data=investment_and_foreign_IO) + 
  geom_histogram(stat = 'identity') +
  geom_point(aes(y=investment,color=country))

grid.arrange(inflow, outflow)
ggsave('us_china_inflow_outflow_comparison.png')


ggplot(aes(x = year, y=foreign_inflow, fill=country), data=investment_and_foreign_IO) + 
  geom_histogram(stat = 'identity') +
  geom_point(aes(y=investment,color=country))

ggplot(aes(x = year, y=foreign_outflow, fill=country), data=investment_and_foreign_IO) + 
  geom_histogram(stat = 'identity') +
  geom_point(aes(y=investment,color=country))

ggplot(aes(x=year, y=net_in, fill=country), data=investment_and_foreign_IO) + geom_histogram(stat = 'identity') 

ggsave('us_china_foreign_inv_net_inflow.png')

```

# Foreign investment inflow in China started 1982.  By 1993, the inflow amount increase 10 fold.  China has been having positive net foreign direct investment inflow since 1982 while US has periods of alternating positive and negative net inflow.  China's foreign direct investment netinflow reached its peak in 1994, has since been taking a downtrend although is still positive.

```{r - agriculture, industry, export, foreign direct investment}
colnames(agriculture)[1] <- 'country'
colnames(industry)[1] <- 'country'
colnames(export)[1] <- 'country'


agriculture <- agriculture %>%
  gather('year', 'agriculture', 2:53)
agriculture$year <- gsub('X', '', agriculture$year)

industry <- industry %>%
  gather('year', 'industry', 2:53)
industry <- gsub('X', '', industry$year)

export <- export %>%
  gather('year', 'export', 2:53)
export$year <- gsub('X', '', export$year)


```

```{r}
us_factors_p_of_GDP <- left_join(subset(agriculture, country=='United State'), subset(industry,  country=='United State'), by = 'year')
us_factors_p_of_GDP <- left_join(us_factors_p_of_GDP, subset(foreign_direct_inv,  country=='United State'))
us_factors_p_of_GDP <- left_join(us_factors_p_of_GDP, subset(export,  country=='United State'))

```




