Lesson 3 - Problem Set
========================================================
***
###  Load Diamond data

***
```{r }
getwd()

library(ggplot2)
data("diamonds")
summary(diamonds)

?diamonds
diamonds$color

```

***
###Create a histogram of the price of all the diamonds in the diamond data set.

***
```{r histogram of diamonds price}
qplot(x=price, data=diamonds)

qplot(x=price, data=diamonds) + scale_x_log10()
```
***
### Diamond Counts - number of diamonds with different price range

***

```{r number of diamonds with different price range}
length(which(diamonds$price < 500))
length(which(diamonds$price < 250))
length(which(diamonds$price >= 15000))

sum(diamonds$price < 500)
```

***
### Cheap Diamonds - Explore the largest peak in the price histogram created earlier.

***

```{r}
qplot(x=price, data=subset(diamonds, !is.na(price)),   
      binwidth = 50, color=I('black'), fill=I('purple')) + 
      scale_x_continuous(lim=c(0, 3000), breaks=seq(0, 3000, 100))
ggsave('pricehistogram.png')

summary(diamonds$price)
    
```

***
### Price by cut histograms

***

```{r Price by cut histograms}
qplot(x=cut, data=subset(diamonds, !is.na(price)))

```


***
### Price by cut histograms

***

```{r Price by cut histograms}
qplot(x=cut, data=subset(diamonds, !is.na(price)))

```



***
### Price by Cut

***

```{r price by cut}
by(diamonds$price, diamonds$cut, summary, digits = max(getOption('digits')))

```

***
### Scales and Multiple Histogram

***

```{r Scales and Multiple Histogram}
qplot(x = price, data = diamonds) + facet_wrap(~cut, scales='free_y')


```
```{r}
qplot(x = price, data = diamonds) + facet_grid(~cut)
```

***
### Price per carat by cut

***

```{r Price per carat by cut}
qplot(x = price/carat, data = diamonds,  color=I('black'),  
        fill=I('purple'), xlab='Price per Carat') + 
        scale_x_log10() + 
        facet_wrap(~cut, scales='free_y')

```

***
### Price Box Plot

***

```{r Price Box Plot - by cut}
qplot(x=cut, y=price, data=subset(diamonds, !is.na(price)),
      color=I('black'), fill=color, geom='boxplot') +
      scale_y_continuous(lim=c(1000, 3000))
ggsave('priceHistogramByCut.png')

```

```{r Price Box Plot - by clarity}
qplot(x=clarity, y=price, data=subset(diamonds, !is.na(price)),
      color=I('black'), fill=color, geom='boxplot') +
      scale_y_continuous(lim=c(1000, 17500))
ggsave('priceHistogramByClarity.png')
```

```{r Price Box Plot - by color}
qplot(x=color, y=price, data=subset(diamonds, !is.na(price)),
      color=I('black'), fill=color, geom='boxplot') +
      coord_cartesian(ylim=c(1000, 7500))
ggsave('priceHistogramByColor.png')
```

***
### Interquartile Range - IQR

***

```{r Interquartile Range - IQR}
by(diamonds$price, diamonds$color, summary)
by(diamonds$price, diamonds$clarity, summary)
by(diamonds$price, diamonds$cut, summary)

?diamonds
```

***
### Price Per Carat Box Plot

***

```{r Price Per Carat Box Plot - by cut}
qplot(x=cut, y=price/carat, data=subset(diamonds, !is.na(price)),
      color=I('black'), fill=color, geom='boxplot') +
      scale_y_continuous(lim=c(1000, 3500))
ggsave('pricePerCaratHistogramByCut.png')

```

```{r Price Per Carat Box Plot - by clarity}
qplot(x=clarity, y=price/carat, data=subset(diamonds, !is.na(price)),
      color=I('black'), fill=color, geom='boxplot') +
      scale_y_continuous(lim=c(1000, 17500))
ggsave('pricePerCaratHistogramByClarity.png')
```

```{r Price Per Carat Box Plot - by color}
qplot(x=color, y=price/carat, data=subset(diamonds, !is.na(price)),
      color=I('black'), fill=color, geom='boxplot') +
      coord_cartesian(ylim=c(1500, 6000))
ggsave('pricePerCaratHistogramByColor.png')

```

```{r}
by(diamonds$price/diamonds$carat, diamonds$color, summary)
by(diamonds$price/diamonds$carat, diamonds$clarity, summary)
by(diamonds$price/diamonds$carat, diamonds$cut, summary)
```


***
### Carat Frequency Polygon

***

```{r Carat Frequency Polygon}
qplot(x=carat, data=diamonds, binwidth=0.3, geom='freqpoly') +
      scale_x_continuous(lim=c(0, 3), breaks=seq(0, 3, 0.1))

```

```{r}
countByCarat <- table(diamonds$carat)
countByCarat[countByCarat > 2000] 
```
***
Result from above
 0.3 0.31 1.01 
2604 2249 2242 
***
***
### Gapminder Data - women labor participation and unployment rate

***

```{r Gapminder Data - women labor participation and unployment rate}

labor <- read.csv("./f_25_54_labour_to_pop_percent/data.csv", head=TRUE, sep=',')
umploy <- read.csv("./f_25_54_unemploy/data.csv", head=TRUE, sep=',')

```

```{r - add library}
install.packages('dplyr')
library(dplyr)
install.packages('tidyr')
library(tidyr)

library(ggplot2)
```

```{r - extract us data and combine}
us_participation <- filter(labor, country == 'United States')
participationRateByYear <- gather(us_participation, 'year', 'labor', 2:30)

us_unemploy <- filter(umploy, country =='United States')
unemployRateByYear <- gather(us_unemploy, 'year', 'unemploy', 2:27)

us_rates <- left_join(participationRateByYear, unemployRateByYear, by = 'year')

us_rates %>% mutate(ratio = unemploy / labor)

View(us_rates)
```

```{r - graphs}
qplot(x=year, y=labor, data=us_rates)
qplot(x=year, y=unemploy, data=us_rates)

ggplot(us_rates, aes(x=year, y = value, color = variable)) + 
    geom_point(aes(y = labor, col = "Labor Participation %")) + 
    geom_point(aes(y = unemploy, col = "Unemployment %")) +
    geom_point(aes(y = unemploy / labor, col='Unemploy / Participation'))
ggsave('f_25_54_unemploy_rate_vs_participation.png')
```
***
Conclusion: Women unemplpoyment rate drop as participation rate increase.  The ratio of unemployment vs. participation stays constant relatively speaking.
***
```{r}
rm(us_participation)
rm(us_unemploy)
rm(us_rates)
rm(labor)
rm(umploy)
rm(participationRateByYear)
rm(p)

```

***
### Explore your friends' Birthdays

***

```{r Explore your friends' Birthdays}
birthdays = read.csv('birthdaysExample.csv')
birthdays <- separate(birthdays, dates, c('month', 'day', 'year'), sep='/')

```

```{r birth month distribution by month}
birthdays$month = as.numeric(birthdays$month)
birthdays$day = as.numeric(birthdays$day)

ggplot(birthdays, aes(x=month, color='black', fill='purple')) + ggtitle('Birth Months Distribution') + geom_bar() + scale_x_discrete('month')
ggsave('birthMonthDistribution.png')

ggplot(birthdays, aes(x=day, color='black', fill='purple')) + ggtitle('Birth Days Distribution') + geom_bar() + scale_x_discrete('day')
ggsave('birthDaysDistribution.png')
```


