---
title: "White Wine Quality Exploration"
author: "Li Wang"
date: "12/1/2016"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#install.packages("ggplot2", dependencies = T) 
require(ggplot2)
#install.packages("knitr", dependencies = T)
#install.packages("dplyr", dependencies = T)
library(dplyr)
#install.packages('tidyr')
library(tidyr)
#install.packages("gridExtra")
library("gridExtra")
library(reshape2)
#install.packages("psych")
library(psych)
#install.packages("corrplot")
library(corrplot)
library(memisc)
```
## 1. Abstract

The purpose of this project is to determine what physicochemical properties affect white wine quality through exploratory data analysis of a dataset containing attributes for approximately 5,000 white variants of the Portuguese "Vinho Verde" wine.  

For more details about the wine, consult: [Cortez et al., 2009](http://www.vinhoverde.pt/en/ or the reference).  Due to privacy and logistic issues, only physicochemical (inputs) and sensory (the output) variables are available (e.g. there is no data about grape types, wine brand, wine selling price, etc.).

## 2. Dataset

This dataset consists of 12 variables, with almost 5,000 observations.
Input variables (based on physicochemical tests):
1 - fixed acidity (tartaric acid - g / dm^3)
2 - volatile acidity (acetic acid - g / dm^3)
3 - citric acid (g / dm^3)
4 - residual sugar (g / dm^3)
5 - chlorides (sodium chloride - g / dm^3
6 - free sulfur dioxide (mg / dm^3)
7 - total sulfur dioxide (mg / dm^3)
8 - density (g / cm^3)
9 - pH
10 - sulphates (potassium sulphate - g / dm3)
11 - alcohol (% by volume)
Output variable (based on sensory data): 
12 - quality (score between 0 and 10)

## 3. Exploratory Analysis

### 3.1 Dataset Preparation and Transformation

```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
white <- read.csv('wineQualityWhites.csv', sep = ',', header = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
dim(white)
str(white)
summary(white)
```

```{r}
# Define original variable names
ORIGINAL <- colnames(white)
# Define independent variable names
INDEPENDENT <- colnames(white)[1:11]

#corrplot.mixed(cor(white[, ORIGINAL]))
```

```{r}
white$quality.f2 <- ifelse(white$quality == 3 | 
                          white$quality == 4 | 
                          white$quality == 5, 
                        "bad", 
                        ifelse(white$quality == 6, 
                               "normal",
                               "good"))
white$quality.f2 <- factor(white$quality.f2, levels=c("bad", "normal", "good"))
table(white$quality.f2)

```
Create a new variable quality.f2 which has fewer quality levels and see if it can provide new insights.

```{r include=FALSE}
#ggplot(aes(x=density.bucket), data=white) + geom_histogram(stat = 'count')
white$density.bucket <- cut(white$density, breaks = c(0.985, 0.990, 0.992, 0.994, 0.996, 0.998, 1.0))
table(white$density.bucket)
```

```{r include=FALSE}
length(which(white$quality == 0))
length(which(white$quality >= 8))
length(which(white$quality <= 4))
#There is no wine in the dataset without  a quality rating.  There are about equal amount of best quality wines and worst quality ones.
```


### 3.2 Univariate Plots Section

```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE }
#p1 <- ggplot(aes(x=fixed.acidity), data=white) + geom_histogram()
#p2 <- ggplot(aes(x=volatile.acidity), data=white) + geom_histogram()
#p3 <- ggplot(aes(x=citric.acid), data=white) + geom_histogram()
#p4 <- ggplot(aes(x=residual.sugar), data=white) + geom_histogram()
#p5 <- ggplot(aes(x=chlorides), data=white) + geom_histogram()
#p6 <- ggplot(aes(x=free.sulfur.dioxide), data=white) + geom_histogram()
#p7 <- ggplot(aes(x=total.sulfur.dioxide), data=white) + geom_histogram()
#p8 <- ggplot(aes(x=density), data=white) + geom_histogram()
#p9 <- ggplot(aes(x=pH), data=white) + geom_histogram()
#p10 <- ggplot(aes(x=sulphates), data=white) + geom_histogram()
#p11 <- ggplot(aes(x=alcohol), data=white) + geom_histogram()
#p12 <- ggplot(aes(x=quality), data=white) + geom_histogram()
#grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol = 3)

```

```{r}
ggplot(ww, aes_string(x=var)) + 
    geom_density(fill='gray') +
    geom_vline(aes(xintercept=mean(ww[, var])), color='blue', size=1) +
    geom_vline(aes(xintercept=median(ww[, var])), color='red', size=1) +
    geom_vline(aes(xintercept=quantile(ww[, var], 0.25)), 
               linetype='dashed', size=0.5) + 
    geom_vline(aes(xintercept=quantile(ww[, var], 0.75)), 
               linetype='dashed', size=0.5)
```


```{r echo=FALSE, message=FALSE, warning=FALSE }
#ggplot(aes(x=quality), data=white) + geom_histogram(stat = 'identity') 
ggplot(aes(x=quality, fill=quality), data=white) + 
  geom_histogram(aes(fill=..count..)) +
  scale_fill_gradient("Count", low = "yellow", high = "red")
```
Most wines are grade 5, 6, 7.


```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE }
#### fixed.Acidity
ggplot(aes(x=fixed.acidity), data=white) + geom_histogram(binwidth = 0.1, fill="blue") 
summary(white$fixed.acidity)
```
Fixed Acidity refers to most acids involved with wine or fixed or nonvolatile (do not evaporate readily).  Most fall between 6 and 7.3 g per cube dm.  

```{r echo=FALSE, message=FALSE, warning=FALSE }
pb_facid <- ggplot(aes(x=fixed.acidity), data=subset(white, quality >= 8)) + geom_histogram(bins=40, fill='#6666FF') + xlim(3.5, 10) + 
  ggtitle('Best Quality Wine')
pw_facid<- ggplot(aes(x=fixed.acidity), data=subset(white, quality <= 4)) + geom_histogram(bins=40, fill='#0099ff') + xlim(3.5, 10) + ggtitle('Worst Quality Wine')
```
The level of fixed acidity in most best quality wine fall within narrow range of 5 to 8 g / dm^3, while the level for worst quality wine fall within 5 to 10 g / dm^3.


```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE }
ggplot(aes(x=volatile.acidity), data=white) + geom_histogram(fill='#660099') + xlim(0, 0.9)
summary(white$volatile.acidity)
```
Volatile Acidity amount skew to the lower end because too high of its levels can lead to an unpleasant, vinegar taste.  Median level is 0.26 g per cube dm.  I wonder if volatile acidity amount is related to wine quality.


```{r echo=FALSE, message=FALSE, warning=FALSE }
pb_vacid <- ggplot(aes(x=volatile.acidity), data=subset(white, quality >= 8)) + geom_histogram(bins=60, fill='#330066') + 
  xlim(0.1, 0.75) + ggtitle('Best Quality Wine')
pw_vacid<- ggplot(aes(x=volatile.acidity), data=subset(white, quality <= 4)) + geom_histogram(bins=60, fill='#9933ff') + 
  xlim(0.1, 0.75) + ggtitle('Worst Quality Wine')

```
There is no distribution difference for volatile acidity amount between best quality wines or worse ones. Best quality wines tend to have less than 0.5 g / dm^3 acetic acid.


```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE }
ggplot(aes(x=citric.acid), data=white) + geom_histogram(fill='#66cc33') + 
  xlim(0, 0.75)
summary(white$citric.acid)
```
Citric Acid found in small quantities, can add 'freshness' and flavor to wines. Most white wine contain 0.2 to 0.4 per cube dm.  


```{r echo=FALSE, message=FALSE, warning=FALSE }
pb_cacid <- ggplot(aes(x=citric.acid), data=subset(white, quality >= 8)) + geom_histogram(bins=40, fill='#339900') + xlim(0, 0.75) + ggtitle('Best Quality Wine')
pw_cacid<- ggplot(aes(x=citric.acid), data=subset(white, quality <= 4)) + geom_histogram(bins=40, fill='#66cc33') + xlim(0, 0.75) + ggtitle('Worst Quality Wine')

```
Best quality wines tend to have consitant citric acid amout, falling within a narrow range of 0.22 to 0.5 g / dm^3.


##### Residual Sugar
```{r echo=FALSE, message=FALSE, warning=FALSE }
ggplot(aes(x=residual.sugar), data=white) + geom_histogram(fill='#990000') + xlim(0, 20)
 
summary(white$residual.sugar)
```
I trimed outliner for the plot since majority wine contains less than 20 g per cube dm.  There is a large count of wine variants contain less than 2 g per cube dm.  I wonder what kind of qulity of those wine.


```{r echo=FALSE, message=FALSE, warning=FALSE }
pb_sugar <- ggplot(aes(x=residual.sugar), data=subset(white, quality >= 8)) + geom_histogram(bins=60, fill='#660000') + 
  xlim(0, 17) + ggtitle('Best Quality Wine')
pw_sugar<- ggplot(aes(x=residual.sugar), data=subset(white, quality <= 4)) + geom_histogram(bins=60, fill='#993366') + 
  xlim(0, 17) + ggtitle('Worst Quality Wine')


```
There is no distribution difference for residual sugar amount between best quality wines or worse ones, although more variants of worst quality wines tend to have low level residual sugar.


#### Chlorides
```{r echo=FALSE, message=FALSE, warning=FALSE }
ggplot(aes(x=chlorides), data=white) + geom_histogram(fill='#339966') + xlim(0, 0.1)
summary(white$chlorides)
```
chlorides is the amount of salt in the wine. Median amount is 0.043 g per cube dm. 


```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
pb_chlorides <- ggplot(aes(x=chlorides), data=subset(white, quality >= 8)) + geom_histogram(bins=60, fill='#006633') + xlim(0, 0.1) + ggtitle('Best Quality Wine')
pw_chlorides<- ggplot(aes(x=chlorides), data=subset(white, quality <= 4)) + geom_histogram(bins=60, fill='#339966') + xlim(0, 0.1) + ggtitle('Worst Quality Wine')


```
There is no different of Chloriades distribution between Best Quality Wine and Worst Quality Wine.


##### Sulfur Dioxide
```{r echo=FALSE, message=FALSE, warning=FALSE }
ggplot(aes(x=free.sulfur.dioxide), data=white) + geom_histogram(fill='#ffcc00') + xlim(0, 100)
#ggplot(aes(x=free.sulfur.dioxide), data=white) + geom_histogram() + scale_x_log10()
summary(white$free.sulfur.dioxide)
```
Free Sulfur Dioxide prevents microbial growth and the oxidation of wine. Its amount ranges from 2 to 289 but most fall under 62 mg per cube dm.


```{r echo=FALSE, message=FALSE, warning=FALSE }
pb_fso2 <- ggplot(aes(x=free.sulfur.dioxide), data=subset(white, quality >= 8)) + geom_histogram(bins=30, fill='#cc9900') + scale_x_log10(limits=c(5, 100)) + ggtitle('Best Quality Wine')
pw_fso2<- ggplot(aes(x=free.sulfur.dioxide), data=subset(white, quality <= 4)) + geom_histogram(bins=30, fill='#ffcc00') + scale_x_log10(limits=c(5, 100)) + ggtitle('Worst Quality Wine')


```
The amount of free sulfur dioxide in Worst quality wine skewd to lower level. After log transformation, its distribution for worst quality wines shows bimodal while that for best quality wine stays in a higher range.


```{r echo=FALSE, message=FALSE, warning=FALSE }
ggplot(aes(x=total.sulfur.dioxide), data=white) + geom_histogram(fill='#cccc00') + xlim(0, 265)
summary(white$total.sulfur.dioxide)
```
Total Sulfur Dioxide (S02) in low concentrations is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine. Most wines contain less than 260 mg per cube dm.


```{r echo=FALSE, message=FALSE, warning=FALSE }
pb_tso2 <- ggplot(aes(x=total.sulfur.dioxide), data=subset(white, quality >= 8)) + geom_histogram(bins=40, fill='#999900') + xlim(0, 250) + ggtitle('Best Quality Wine')
pw_tso2<- ggplot(aes(x=total.sulfur.dioxide), data=subset(white, quality <= 4)) + geom_histogram(bins=40, fill='#cccc00') + xlim(0, 250) + ggtitle('Worst Quality Wine')


```
Total sulfur dioxide amont in most best quality wine falls in a narrow range of 70 to 190, while that in most worst quality wine falls in the range of 5 to 250.


##### Density
```{r echo=FALSE, message=FALSE, warning=FALSE }
ggplot(aes(x=density), data=white) + geom_histogram(bins = 50, fill='#0000cc')
summary(white$density)
```
Density of water depends on the percent alcohol and sugar content.  


```{r echo=FALSE, message=FALSE, warning=FALSE }
pb_density <- ggplot(aes(x=density), data=subset(white, quality >= 8)) + geom_histogram(bins=40, fill='#3333cc')  + ggtitle('Best Quality Wine')
pw_density<- ggplot(aes(x=density), data=subset(white, quality <= 4)) + geom_histogram(bins=40, fill='#0000ff')  + ggtitle('Worst Quality Wine')


```
Most best quality wines density distribution has a narrow range.


##### pH
```{r echo=FALSE, message=FALSE, warning=FALSE }
ggplot(aes(x=pH), data=white) + geom_histogram()
summary(white$pH)
```
pH describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-3.4 on the pH scale


```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
pb_ph <- ggplot(aes(x=pH), data=subset(white, quality >= 8)) + geom_histogram(bins=90) + xlim(2.75, 3.75)  + ggtitle('Best Quality Wine')
pw_ph<- ggplot(aes(x=pH), data=subset(white, quality <= 4)) + geom_histogram(bins=90)  + xlim(2.75, 3.75)  + ggtitle('Worst Quality Wine')


```
There is no different of pH distribution between Best Quality Wine and Worst Quality Wine.


##### Sulphates
```{r echo=FALSE, message=FALSE, warning=FALSE }
ggplot(aes(x=sulphates), data=white) + geom_histogram(bins=50, fill='#996633')
summary(white$sulphates)
```
Sulphates is a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant.  Sulphates additive skew to the right lower end of the distribution range.


```{r echo=FALSE, message=FALSE, warning=FALSE }
pb_sulphates <- ggplot(aes(x=sulphates), data=subset(white, quality >= 8)) + geom_histogram(bins=40, fill='#663300') + xlim(0.15, 1.0)  + ggtitle('Best Quality Wine')
pw_sulphates<- ggplot(aes(x=sulphates), data=subset(white, quality <= 4)) + geom_histogram(bins=40, fill='#996633') + xlim(0.15, 1.0)  + ggtitle('Worst Quality Wine')


```
Sulphate level distribution has not much difference between best quality wines and worse ones.


##### Alcohol
```{r echo=FALSE, message=FALSE, warning=FALSE }
ggplot(aes(x=alcohol), data=white) + geom_histogram(bins = 60, fill='#660066') 
#p_alcohol_log <- ggplot(aes(x=alcohol), data=white) + geom_histogram(bins = 60) + scale_x_log10() + ggtitle('With log transformaiton')
#grid.arrange(p_alcohol, p_alcohol_log)
summary(white$alcohol)
```
Alcohol level skews to the lower end of the distribution range.  


```{r echo=FALSE, message=FALSE, warning=FALSE }
pb_alcohol <- ggplot(aes(x=alcohol), data=subset(white, quality >= 8)) + geom_histogram(bins=60, fill='#660066') +  
  xlim(7.5, 14) + ggtitle('Best Quality Wine')
pw_alcohol<- ggplot(aes(x=alcohol), data=subset(white, quality <= 4)) + geom_histogram(bins=60, fill='#993399') +  
  xlim(7.5, 14) + ggtitle('Worst Quality Wine')


```
Most best quality wine tends to have higher leve of alcohol range from 10.5 to 13.5 % by volume, while most worst quality wine alcohol level range from 8 to 12 % by volume with bimodal distribution.




```{r}
grid.arrange(pb_alcohol, pw_alcohol, ncol=1)
grid.arrange(pb_density, pw_density, ncol=1)
grid.arrange(pb_sugar, pw_sugar, ncol=1)
grid.arrange(pb_chlorides, pw_chlorides, ncol=1)
grid.arrange(pb_tso2, pw_tso2, ncol=1)
grid.arrange(pb_fso2, pw_fso2, ncol=1)
grid.arrange(pb_facid, pw_facid, ncol=1)
grid.arrange(pb_vacid, pw_vacid, ncol=1)
grid.arrange(pb_cacid, pw_cacid, ncol=1)
grid.arrange(pb_ph, pw_ph, ncol=1)
grid.arrange(pb_sulphates, pw_sulphates, ncol=1)


```


### 3.3 Univariate Analysis

Structure of dataset:

There are 4898 white wine observations in the dataset with 12 features.  The variables quality is ordered factor variables with the following levels.

(worst) 1 —————-> (best) 12

Other Observations:

The main features in the data set are density, alcohol and quality. I’d like to determine which features are best for predicting the quality of wine. I suspect acidity and some combination of the other variables can be used to build a predictive model to determine quality of wine.

Fixed acidity, volatile acidity, residual sugar, chlorides, total sulfur.dioxide could also contribute to the quality of wine. 

There is high count of centain low level residual sugar of around 2.


### 3.4 Bivariate Plots Section

```{r}
white2 <- white[c(-1, -4, -7, -11)]

#cor(white2)
lowerCor(white2[sapply(white2, is.numeric)])
# pairs.panels(white)
pairs.panels(white2)
ggsave('pairs_panels_whites.png')
```
Based on coorelation study, citric acid, free SCO2, sulphate are not correlated to quality, fixed acidity, resisual sugar, pH are weakly correlated to quality.  My study will concentrate on how volatile acidity, chloride, total SCO2, alcohol, density would affect wine quality.


#### Quality

```{r include=FALSE}
cor(white$alcohol, white$quality)^2
cor(white$density, white$quality)^2
cor(white$chlorides, white$quality)^2
cor(white$volatile.acidity, white$quality)^2
cor(white$total.sulfur.dioxide, white$quality)^2
cor(white$fixed.acidity, white$quality)^2
cor(white$residual.sugar, white$quality)^2
cor(white$pH, white$quality)^2
cor(white$sulphates, white$quality)^2
cor(white$citric.acid, white$quality)^2
cor(white$free.sulfur.dioxide, white$quality)^2
```

```{r}
# factors vs. quality
a_q <- ggplot(aes(y=alcohol, x=quality, group=quality), data=white) + geom_boxplot()
ch_q <- ggplot(aes(y=chlorides, x=quality, group=quality), data=white) + geom_boxplot()
t_q <- ggplot(aes(y=total.sulfur.dioxide, x=quality, group=quality), data=white) + geom_boxplot()
d_q <- ggplot(aes(y=density, x=quality, group=quality), data=white) + geom_boxplot()
rs_q <- ggplot(aes(y=residual.sugar, x=quality, group=quality), data=white) + geom_boxplot()
fr_q <- ggplot(aes(y=free.sulfur.dioxide, x=quality, group=quality), data=white) + geom_boxplot()
grid.arrange(a_q, ch_q, t_q, d_q, rs_q, fr_q, ncol=3 )
```


```{r}
a_qf2 <- ggplot(white, aes(alcohol, fill=quality.f2)) +
  geom_density() +
  xlim(8, 14) +
  ylab('')

d_qf2 <- ggplot(white, aes(x=density, fill=quality.f2)) + 
  geom_density(alpha=0.3) +
  xlim(0.986, 1.005) +
  ylab('')

grid.arrange(a_qf2, d_qf2)
```


```{r include=FALSE}
by(white$alcohol, white$quality, summary)
alcohol.lm = lm(alcohol ~ quality, data=white)
summary(alcohol.lm)
```
Best quality wines have highest median alcohol level.

```{r include=FALSE}
by(white$density, white$quality, summary)
density.lm = lm(density ~ quality, data=white)
summary(density.lm)
```
Best quality wines have lowest median density and smallest range.

```{r include=FALSE}
sugar.lm = lm(residual.sugar ~ quality, data=white)
summary(sugar.lm)
```

```{r include=FALSE}
by(white$chlorides, white$quality, summary)
chlorides.lm = lm(chlorides ~ quality, data=white)
summary(chlorides.lm)
```
Median Chlorides amount for best quality wines is slightly lower than other grade of wines.  


#### Alcohol

```{r include=FALSE}
cor(white$density, white$alcohol)^2
cor(white$residual.sugar, white$alcohol)^2
cor(white$total.sulfur.dioxide, white$alcohol)^2
cor(white$chlorides, white$alcohol)^2
cor(white$fixed.acidity, white$alcohol)^2
cor(white$pH, white$alcohol)^2
```

```{r include=FALSE}
coef(lm(density ~ alcohol, data=white))
coef(lm(residual.sugar ~ alcohol, data=white))
coef(lm(total.sulfur.dioxide ~ alcohol, data=white))
coef(lm(chlorides ~ alcohol, data=white))
```

```{r}
# factors vs. alcohol
d_a <- ggplot(aes(x=alcohol, y=density, colour=density), data=white) + 
  geom_point() + scale_colour_gradientn(colours=rainbow(4)) +
  geom_abline(intercept = 1.014, slope = -0.0019)

rs_a <- ggplot(aes(x=alcohol, y=residual.sugar, colour=residual.sugar), data=white) + scale_colour_gradientn(colours=rainbow(4)) +
  geom_point(alpha=1/10) + ylim(0, 40) + 
  geom_abline(intercept = 25.9, slope = -1.86)

t_a <- ggplot(aes(x=alcohol, y=total.sulfur.dioxide, color=total.sulfur.dioxide), data=white) + 
  geom_point() + scale_colour_gradientn(colours=rainbow(4)) +
  geom_abline(intercept = 301.4, slope = -15.5)

c_a <- ggplot(aes(x=alcohol, y=chlorides, color=chlorides), data=white) + 
  geom_point() + scale_colour_gradientn(colours=rainbow(4)) +
  geom_abline(intercept = 0.11, slope = -0.006)

grid.arrange(d_a, rs_a, t_a, c_a, ncol=2)
```
Alcohol is strongly correlated to density, and weakly correlated to residual sugar, total sulfur dioxide, and chlorides.


#### Density

```{r include=FALSE}
cor(white$residual.sugar, white$density)^2
cor(white$alcohol, white$density)^2
cor(white$total.sulfur.dioxide, white$density)^2
cor(white$free.sulfur.dioxide, white$density)^2
cor(white$fixed.acidity, white$density)^2
cor(white$chlorides, white$density)^2
cor(white$volatile.acidity, white$density)^2
```

```{r include=FALSE}
coef(lm(residual.sugar ~ density, data=white))
coef(lm(total.sulfur.dioxide ~ density, data=white))
coef(lm(chlorides ~ density, data=white))
coef(lm(free.sulfur.dioxide ~ density, data=white))
```

```{r}
# factors vs. density
rs_d <- ggplot(aes(y=residual.sugar, x=density, colour=residual.sugar), data=white) + 
  geom_point() + xlim(0.985, 1.02) +
  geom_abline(intercept = -1408, slope = 1423)

t_d <- ggplot(aes(y=total.sulfur.dioxide, x=density, colour= total.sulfur.dioxide), data=white) + 
  geom_point() + xlim(0.985, 1.02) + 
  geom_abline(intercept = -7346, slope = 7529)

c_d <- ggplot(aes(y=chlorides, x=density, colour=chlorides), data=white) + 
  geom_point() + xlim(0.985, 1.02) + 
  geom_abline(intercept = -1.81, slope = 1.88)

f_d <- ggplot(aes(y=free.sulfur.dioxide, x=density, colour=free.sulfur.dioxide), data=white) + 
  geom_point() + xlim(0.985, 1.02) + 
  geom_abline(intercept = -1628, slope = 1672)

grid.arrange(rs_d, t_d, c_d, f_d)
```
Denisty is strongly correlated to residual sugar, and weakly correlated to chlorides, total sulfur dioxide, free sulfur dioxide.


#### Residual Sugar
```{r include=FALSE}
coef(lm(density ~ residual.sugar, data=white))
coef(lm(alcohol ~ residual.sugar, data=white))
coef(lm(total.sulfur.dioxide ~ residual.sugar, data=white))
```

```{r}
# factors vs. residual.sugar
rs_d <-ggplot(aes(x=residual.sugar, y=density), data=white) + 
  geom_point(alpha=1/10) +
  geom_abline(intercept = 0.99, slope = 0.0005, color='purple')

rs_a <- ggplot(aes(x=residual.sugar, y=alcohol), data=white) + 
  geom_point(alpha=1/10) + ylim(0, 40) + 
  geom_abline(intercept = 11.2, slope = -0.1, color='orange')

rs_t <- ggplot(aes(x=residual.sugar, y=total.sulfur.dioxide), data=white) + 
  geom_point(alpha=1/10) + 
  geom_abline(intercept = 116.8, slope = 3.4, color='tan')

rs_ph <- ggplot(aes(x=residual.sugar, y=pH), data=white) + geom_point(alpha=1/10)

grid.arrange(rs_d, rs_a, rs_t, rs_ph, ncol=2)

```
Residual sugar level is strongly positively correlated to density, weakly negatively correlated to alcohol and posivitively correlated to total sulfur dioxide.  Residual sugar is not correlated to pH.


#### Total Sulfur Dioxide

```{r include=FALSE}
coef(lm(density ~ total.sulfur.dioxide, data=white))
coef(lm(alcohol ~ total.sulfur.dioxide, data=white))

#total_so2.lm = lm(total.sulfur.dioxide ~ quality, data=white)
#summary(total_so2.lm)
```

```{r}
# factors vs. total.sulfur.dioxide
d_t <- ggplot(aes(x=total.sulfur.dioxide, y=density, color=density), data=white) + 
  geom_point(alpha=1/10, position = 'jitter') +
  scale_colour_gradientn(colours=rainbow(4)) +
  geom_abline(intercept = 0.99, slope = 0.00003, color='blue')

a_t <- ggplot(aes(x=total.sulfur.dioxide, y=alcohol, color=alcohol), data=white) + 
  geom_point(alpha=1/5) +
  scale_colour_gradientn(colours=rainbow(4)) +
  geom_abline(intercept = 12.31301, slope = -0.013, color='purple')

grid.arrange(d_t, a_t, ncol=1)
```
Total sulfur dioxide is strongly correlated to alcohol.


#### Chlorides

```{r include=FALSE}
coef(lm(density ~ chlorides, data=white))
coef(lm(alcohol ~ chlorides, data=white))
coef(lm(total.sulfur.dioxide ~ chlorides, data=white))

```

```{r}
# factors vs. chlorides
d_ch <- ggplot(aes(x=chlorides, y=density, color=density), data=white) + 
  geom_point(alpha=1/5) +
  scale_colour_gradientn(colours=rainbow(6)) +
  geom_abline(intercept = 0.992, slope = 0.035, color='blue')

a_ch <- ggplot(aes(x=chlorides, y=alcohol, color=alcohol), data=white) + 
  geom_point(alpha=1/5) +
  scale_colour_gradientn(colours=rainbow(6)) +
  geom_abline(intercept = 11.44, slope = -20.288, color='red')

t_ch <- ggplot(aes(x=chlorides, y=total.sulfur.dioxide, color=total.sulfur.dioxide), data=white) + 
  geom_point(alpha=1/5) +
  scale_colour_gradientn(colours=rainbow(6)) +
  geom_abline(intercept = 120.7, slope = 386.9, color='orange')

grid.arrange(d_ch, a_ch, t_ch, ncol=1)

```
Chlorides is positively correlated to density although not a strong relationship.  Chlorides is slightly correlated to total.sulfur.dioxide.


#### Fixed acidity

```{r include=FALSE}
coef(lm(density ~ fixed.acidity, data=white))
coef(lm(pH ~ fixed.acidity, data=white))
```

```{r}
# factors vs. fixed acidity
d_fa<- ggplot(aes(x=fixed.acidity, y=density, color=density), data=white) + 
  geom_point(alpha=1/10, position = 'jitter')  +
  geom_abline(intercept = 0.98, slope = 0, color='red')

ph_fa <- ggplot(aes(x=fixed.acidity, y=pH, color=pH), data=white) + 
  geom_point(alpha=1/5) + 
  geom_abline(intercept = 3.7, slope = -0.076, color='orange')

grid.arrange(d_fa, ph_fa, ncol=1)
```
Fixed acidity is negatively correlated to pH and positively correlated to density weakly. 


```{r include=FALSE}
#### pH
cor(white$fixed.acidity, white$pH)^2
cor(white$residual.sugar, white$pH)^2
cor(white$citric.acid, white$pH)^2
cor(white$volatile.acidity, white$pH)^2
#pH is weakly correlated to fixed acidity.
```


### 3.5 Bivariate Analysis

Two more significiant independent variables correlating to quality are alcohol and density. 

In terms of relationships between independent variables, some strong correlations are observed.

0.84 residual.sugar - density
0.78 alcohol - density
0.62 free.sulfur.dioxide - total.sulfur.dioxide
0.53 total.sulfur.dioxide - density
0.45 residual sugar - alcohol
0.36 chloride - alchoho


### 3.6 Multivariate Plots and Analysis 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(y=alcohol, x=density, color=quality), data=white) + 
  geom_point() +
  facet_wrap(~ quality.f2) +
  scale_colour_gradientn(colours=rainbow(6))
```
Within the same range of density, best quality wines have highest level of alcohol.


```{r }
# Colored by simplified quality level
rs.qf2 <- ggplot(white, aes(x=density, y=residual.sugar, color=quality.f2)) +
  geom_point(position='jitter') + 
  xlim(min(white$density), quantile(white$density, prob=0.99)) +
  ylim(min(white$residual.sugar), quantile(white$residual.sugar, prob=0.99)) + 
  scale_color_brewer(type='div')
# Colored by original quality level
rs.qf <- ggplot(white, aes(x=density, y=residual.sugar, color=quality.f2)) +
  geom_point() + 
  scale_color_brewer() +
  xlim(min(white$density), quantile(white$density, prob=0.99)) +
  ylim(min(white$residual.sugar), quantile(white$residual.sugar, prob=0.99))

# Colored by simplified quality level
a.qf2 <- ggplot(white, aes(x=density, y=alcohol, color=quality.f2)) +
  geom_point(position='jitter') + 
  xlim(min(white$density), quantile(white$density, prob=0.99)) +
  ylim(min(white$alcohol), quantile(white$alcohol, prob=0.99))  + 
  scale_color_brewer(type='div')
# Colored by original quality level
a.qf <- ggplot(white, aes(x=density, y=alcohol, color=quality.f2)) +
  geom_point() + 
  scale_color_brewer() +
  xlim(min(white$density), quantile(white$density, prob=0.99)) +
  ylim(min(white$alcohol), quantile(white$alcohol, prob=0.99))

# Colored by simplified quality level
t.qf2 <- ggplot(white, aes(x=density, y=total.sulfur.dioxide, color=quality.f2)) +
  geom_point(position='jitter') + 
  xlim(min(white$density), quantile(white$density, prob=0.99)) +
  ylim(min(white$total.sulfur.dioxide), quantile(white$total.sulfur.dioxide, prob=0.99))  + 
  scale_color_brewer(type='div')
# Colored by original quality level
t.qf <- ggplot(white, aes(x=density, y=total.sulfur.dioxide, color=quality.f2)) +
  geom_point() + 
  scale_color_brewer() +
  xlim(min(white$density), quantile(white$density, prob=0.99)) +
  ylim(min(white$total.sulfur.dioxide), quantile(white$total.sulfur.dioxide, prob=0.99))

grid.arrange(rs.qf2, rs.qf, a.qf2, a.qf, t.qf2, t.qf, ncol=2)
```
The above plot resiual sugar, alcohol, and total sulfur dioxide against density seperately.  The left column uses the simplified quality level and the right one uses the original quality level. In both kinds of plots, holding density (mostly from the lower end of density), higher  residual.sugar or alcoho, or total sulfur dioxide seem to have better quality respectively.


```{r include=FALSE }
# similiar plots with density.bucket.  Not very insightful.
# Colored by simplified quality level
rs.dbqf2 <- ggplot(white, aes(x=density.bucket, y=residual.sugar, color=quality.f2)) +
  geom_point(position='jitter')  
# Colored by original quality level
rs.dbqf <- ggplot(white, aes(x=density.bucket, y=residual.sugar, color=quality.f2)) +
  geom_point() + 
  scale_color_brewer(type='seq') 

# Colored by simplified quality level
a.dbqf2 <- ggplot(white, aes(x=density.bucket, y=alcohol, color=quality.f2)) +
  geom_point(position='jitter')    
# Colored by original quality level
a.dbqf <- ggplot(white, aes(x=density.bucket, y=alcohol, color=quality.f2)) +
  geom_point() + 
  scale_color_brewer(type='seq') 

# Colored by simplified quality level
t.dbqf2 <- ggplot(white, aes(x=density.bucket, y=total.sulfur.dioxide, color=quality.f2)) +
  geom_point(position='jitter')  
# Colored by original quality level
t.dbqf <- ggplot(white, aes(x=density.bucket, y=total.sulfur.dioxide, color=quality.f2)) +
  geom_point() + 
  scale_color_brewer(type='seq') 

grid.arrange(rs.dbqf2, rs.dbqf, a.dbqf2, a.dbqf, t.dbqf2, t.dbqf, ncol=2)
```


```{r}
# Colored by simplified quality level
rsa.qf2 <- ggplot(white, aes(x=alcohol, y=residual.sugar, color=quality.f2)) +
  geom_point(position='jitter') + 
  xlim(min(white$alcohol), quantile(white$alcohol, prob=0.99)) +
  ylim(min(white$residual.sugar), quantile(white$residual.sugar, prob=0.99))
# Colored by original quality level
rsa.qf <- ggplot(white, aes(x=alcohol, y=residual.sugar, color=quality.f2)) +
  geom_point() + 
  scale_color_brewer() +
  xlim(min(white$alcohol), quantile(white$alcohol, prob=0.99)) +
  ylim(min(white$residual.sugar), quantile(white$residual.sugar, prob=0.99))

# Colored by simplified quality level
ca.qf2 <- ggplot(white, aes(x=alcohol, y=chlorides, color=quality.f2)) +
  geom_point(position='jitter') + 
  xlim(min(white$alcohol), quantile(white$alcohol, prob=0.99)) +
  ylim(min(white$chlorides), quantile(white$chlorides, prob=0.99))
# Colored by original quality level
ca.qf <- ggplot(white, aes(x=alcohol, y=chlorides, color=quality.f2)) +
  geom_point() + 
  scale_color_brewer() +
  xlim(min(white$alcohol), quantile(white$alcohol, prob=0.99)) +
  ylim(min(white$chlorides), quantile(white$chlorides, prob=0.99))

grid.arrange(rsa.qf2, rsa.qf, ca.qf2, ca.qf, ncol=2)
```
The above plot resiual sugar, chlorides against alcohol seperately.  The left column uses the simplified quality level and the right one uses the original quality level. In both kinds of plots, holding residual.sugar or chlorides respectively, higher alcohol level seem to have better quality.


```{r}
# Colored by simplified quality level
ft.qf2 <- ggplot(white, aes(x=free.sulfur.dioxide, y=total.sulfur.dioxide, color=quality.f2)) +
  geom_point(position='jitter') + 
  xlim(min(white$free.sulfur.dioxide), quantile(white$free.sulfur.dioxide, prob=0.99)) +
  ylim(min(white$total.sulfur.dioxide), quantile(white$total.sulfur.dioxide, prob=0.99))
# Colored by original quality level
ft.qf <- ggplot(white, aes(x=free.sulfur.dioxide, y=total.sulfur.dioxide, color=quality.f2)) +
  geom_point() + 
  scale_color_brewer() +
  xlim(min(white$free.sulfur.dioxide), quantile(white$free.sulfur.dioxide, prob=0.99)) +
  ylim(min(white$total.sulfur.dioxide), quantile(white$total.sulfur.dioxide, prob=0.99))

grid.arrange(ft.qf2, ft.qf, ncol=2)
```
Looking at the above plots, it seems that there are more better quality wine under 150 g dm^3 total sulfur dioxide. 


## 4. Linear Regression

Based on the exploratory analysis in the previous section, there does not seem to be any simple linear relationship between quality and physicochemical properties. If this observation is correct, linear regression model would not perform so well in terms of quality prediction by physicochemical properties.

```{r}
m1 <- lm(I(alcohol) ~ I(quality), data = white)
m2 <- update(m1, ~ . + density)
#m3 <- update(m2, ~ . + residual.sugar)
#m4 <- update(m3, ~ . + total.sulfur.dioxide)
#m5 <- update(m4, ~ . + chlorides)
mtable(m1,m2)
```


```{r}
m1 <- lm(I(density) ~ I(quality), data = white)
m2 <- update(m1, ~ . + alcohol)
#m3 <- update(m2, ~ . + residual.sugar)
#m4 <- update(m3, ~ . + total.sulfur.dioxide)
#m5 <- update(m4, ~ . + chlorides)
mtable(m1,m2)
```
Looking at the statistics summary of two linear model.  It seems that only 20% of variance of quality is explained by alcohol, 10 % by density.  The rest of the result do not make sense.  Linear model is not a suitable approach to predict quality.


## 5.Final Plots and Summary

```{r}
ggplot(aes(x=quality, y=alcohol, group=quality, fill=quality), data=white) + 
  geom_boxplot() +
  ggtitle("Alcohol vs. Quality") + 
  xlab("Quality)") +
  ylab("Alcohol (% by volume)") +
  theme(title=element_text(size=14, face='bold'),
      axis.title=element_text(size=12), 
      axis.text=element_text(size=10),
      legend.text=element_text(size=10))

```
Best quality wines have highest median alcohol level.


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(white, aes(x=density, fill=quality.f2)) + 
  geom_density(alpha=0.3) +
  xlim(quantile(white$density, prob=0.00), 
       quantile(white$density, prob=0.99)) +
  guides(fill=guide_legend(title="Quality", reverse=T)) +
  ggtitle("Distribution of Density per Quality Level") + 
  xlab("Density (g/ml)") +
  ylab("") +
  theme(title=element_text(size=14, face='bold'),
      axis.title=element_text(size=12), 
      axis.text=element_text(size=10),
      legend.text=element_text(size=10))

```
This is a density plot of density. As quality goes up, the center of distribution of density gets smaller.


```{r}
# Colored by simplified quality level
rs.qf2 <- ggplot(white, aes(x=density, y=residual.sugar, color=quality.f2)) +
  geom_point(position='jitter') + 
  scale_color_brewer(type = 'div') +
  xlim(min(white$density), quantile(white$density, prob=0.99)) +
  ylim(min(white$residual.sugar), quantile(white$residual.sugar, prob=0.99)) +
  ggtitle("Residual Sugar vs. Density per Quality Factor") + 
  xlab("Density (g/ml)") +
  ylab("Residual Sugar (g / dm^3)") +
  theme(title=element_text(size=10, face='bold'),
      axis.title=element_text(size=8), 
      axis.text=element_text(size=8),
      legend.text=element_text(size=8))

# Colored by original quality level
rs.qf <- ggplot(white, aes(x=density, y=residual.sugar, color=quality.f2)) +
  geom_point() + 
  scale_color_brewer() +
  xlim(min(white$density), quantile(white$density, prob=0.99)) +
  ylim(min(white$residual.sugar), quantile(white$residual.sugar, prob=0.99)) +
  ggtitle("Residual Sugar vs. Density per Quality Level") + 
  xlab("Density (g/ml)") +
  ylab("Residual Sugar (g / dm^3)") +
  theme(title=element_text(size=10, face='bold'),
      axis.title=element_text(size=8), 
      axis.text=element_text(size=8),
      legend.text=element_text(size=8))

# Colored by simplified quality level
a.qf2 <- ggplot(white, aes(x=density, y=alcohol, color=quality.f2)) +
  geom_point(position='jitter') + 
  scale_color_brewer(type = 'div') +
  xlim(min(white$density), quantile(white$density, prob=0.99)) +
  ylim(min(white$alcohol), quantile(white$alcohol, prob=0.99)) + 
  ggtitle("Alcohol vs. Density per Quality Factor") + 
  xlab("Density (g/ml)") +
  ylab("Alcohol (% by volume)") +
  theme(title=element_text(size=10, face='bold'),
      axis.title=element_text(size=8), 
      axis.text=element_text(size=8),
      legend.text=element_text(size=8))
# Colored by original quality level
a.qf <- ggplot(white, aes(x=density, y=alcohol, color=quality.f2)) +
  geom_point() + 
  scale_color_brewer() +
  xlim(min(white$density), quantile(white$density, prob=0.99)) +
  ylim(min(white$alcohol), quantile(white$alcohol, prob=0.99)) + 
  ggtitle("Alcohol vs. Density per Quality Level") + 
  xlab("Density (g/ml)") +
  ylab("Alcohol (% by volume)") +
  theme(title=element_text(size=10, face='bold'),
      axis.title=element_text(size=8), 
      axis.text=element_text(size=8),
      legend.text=element_text(size=8))

# Colored by simplified quality level
t.qf2 <- ggplot(white, aes(x=density, y=total.sulfur.dioxide, color=quality.f2)) +
  geom_point(position='jitter') + 
  scale_color_brewer(type = 'div') +
  xlim(min(white$density), quantile(white$density, prob=0.99)) +
  ylim(min(white$total.sulfur.dioxide), quantile(white$total.sulfur.dioxide, prob=0.99)) + 
  ggtitle("Total Sulfur Dioxide vs. Density per Quality Factor") + 
  xlab("Density (g/ml)") +
  ylab("Total Sulfur Dioxide (mg / dm^3)") +
  theme(title=element_text(size=10, face='bold'),
      axis.title=element_text(size=8), 
      axis.text=element_text(size=8),
      legend.text=element_text(size=8))
# Colored by original quality level
t.qf <- ggplot(white, aes(x=density, y=total.sulfur.dioxide, color=quality.f2)) +
  geom_point() + 
  scale_color_brewer() +
  xlim(min(white$density), quantile(white$density, prob=0.99)) +
  ylim(min(white$total.sulfur.dioxide), quantile(white$total.sulfur.dioxide, prob=0.99)) + 
  ggtitle("Total Sulfur Dioxide vs. Density per Quality Level") + 
  xlab("Density (g/ml)") +
  ylab("Total Sulfur Dioxide (mg / dm^3)") +
  theme(title=element_text(size=10, face='bold'),
      axis.title=element_text(size=8), 
      axis.text=element_text(size=8),
      legend.text=element_text(size=8))

grid.arrange(rs.qf2, rs.qf, a.qf2, a.qf, t.qf2, t.qf, ncol=2)

```
The above plot resiual sugar, alcohol, and total sulfur dioxide against density seperately.  The left column uses the simplified quality level and the right one uses the original quality level. In both kinds of plots, holding density (mostly from the lower end of density), higher  residual.sugar or alcoho, or total sulfur dioxide seem to have better quality respectively.


## 6. Reflection
Initially I thought I could explore relationships between independent variables and find some significant ones and build a linear model to predict wine quality.  Things didn't turn out like that.  Wine is nothing like vinegar or soy sauce.  It is full of delicateness and subtleness.  Linear regression is not suitable to predict wine quality.  

## 7. Reference
[Creating Effective Plots](https://docs.google.com/document/d/1-f3wM3mJSkoWxDmPjsyRnWvNgM57YUPloucOIl07l4c/pub)
[Colors (ggplot2)](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#palettes-color-brewer)
[My Commonly Done ggplot2 graphs: Part 2](https://www.r-bloggers.com/my-commonly-done-ggplot2-graphs-part-2/)
[Teru Watanabe] (https://rpubs.com/watanabe8760/white-wine)


